Agreement Attraction in Turkish

### 2020-09-01 ongoing_draft/ongoing_paper.tex

---
title             : "Agreement Attraction in Turkish"
shorttitle        : "Agreement Attraction in Turkish"
author: 
  - name          : "Utku Türk"
    affiliation   : "1"
    corresponding : yes    # Define only one corresponding author
    address       : "Postal address"
    email         : "utku.turk@boun.edu.tr"
affiliation:
  - id            : "1"
    institution   : "Boğaziçi University University, Istanbul, Turkey"
authornote: |
  authornote
abstract: |
  We report the results of two speeded acceptability judgment experiments in Turkish. We hypothesized an alternative explanation for agreement attraction effects in Turkish that is based on shallow processing. Our findings contradict our hypothesized form-driven processing strategy and support an account of agreement attraction based on the use of abstract linguistic features, rather than mere form.
keywords          : "keywords"
wordcount         : "X"
bibliography      : ["new-references.bib","r-references.bib"]
floatsintext      : yes
figurelist        : no
tablelist         : no
numbersections    : yes
geometry          : left=25mm, right=25mm, top=25mm, bottom=25mm
#fontsize          : 10pt
footnotelist      : no
linenumbers       : no
mask              : no
draft             : no
documentclass     : "apa6"
classoption       : "doc"
output:
  papaja::apa6_pdf:
    latex_engine: xelatex
    includes:
        in_header: paper_draft_preamble.tex
editor_options: 
  chunk_output_type: console
keep_tex: FALSE
keep_md: FALSE
---

```{r setup, eval =T, include = FALSE, message=FALSE, warning=FALSE}
library("papaja")
library(knitcitations)
library(tidyverse)
library(magrittr)
library(ggplot2)
theme_set(theme_bw())

library(car, warn.conflicts = FALE)
library(MASS)
library(brms)
library(xtable)
library(ggpubr)

library(languageR)
library(tidyverse)
library(gdata)
library(MASS)
library(magrittr)
library(ggplot2)

source("../scripts/misc.R")

```

```{r analysis-preferences, eval =T, message=FALSE, warning=FALSE}
# Seed for random number generation
set.seed(42)
knitr::opts_chunk$set(encoding = 'UTF-8',
                      cache.extra = knitr::rand_seed,
                      echo = FALSE,
                      results = 'asis')
options("citation_format" = "pandoc")

```

# Introduction

It has been observed that speakers often fail to accurately process grammatical dependencies between the parts of a sentence [@GibsonThomas:1999; @PhillipsEtAl:2011]. 
For example, ungrammatical center-embedded structures like  *The patient who the nurse who the clinic had hired met Jack* are found grammatical by native English speakers. 
One such comprehension error is agreement attraction [@Nicoletal:1997; @PearlmutterGarnseyBock:1999; @WagersEtAl:2009]. 
Speakers may erroneously find sentences acceptable in which the verb erroneously agrees with a syntactically unrelated noun phrase (*the attractor*) instead of the subject (*the agreement controller*) as in \@ref(item:wagers2009).

@@

<!-- for some reason I could not add citation in the examples. \hfill \footnotesize{[@WagersEtAl:2009]} -->

These effects were found to be robust in a variety of constructions in different languages, e.g. English , German , Russian, Arabic, Spanish, Armenian, Slovak, Greek, and Turkish [@LagoEtAl2018; @LagoEtAl:2015; @ReifegersteEtAl:2016; @BadeckerKuminiak:2007; @TuckerEtAl:2015; @PaspaliMarinis:2020; @AvetisyanEtAl:2020] using structures like object relative clauses, subject relative clauses, possessive relative clauses, and genitive-possessive noun phrases [@LagoEtAl2018; @WagersEtAl:2009; @LagoEtAl:2015; @FranckColonnaRizzi:2015; @TuckerEtAl:2015; @ParkerEtAl:2015; @Dillon2013a; @Haussler2009]. These effects are found both in grammatical marking of the gender and the number [@BadeckerKuminiak:2007; @Slioussar:2016; @PaspaliMarinis:2020], and attested with various methods including maze tasks, ERP, eye tracking to name a few [@TannerEtAl:2014; @WagersEtAl:2009; @Nicoletal:1997].

One instance of agreement attraction is found in the production processes in which participants were asked to complete sentence fragments [@Bock1991]. Participants used the wrong verb more often when the attractor (*cabinets*) were plural, and the agreement controller (*key*) were singular as in \@ref(item:bock1991).

@@

Agreement attraction is also found in the comprehension studies. In these studies, agreement attraction is found either as a facilitation in reading times or as a effect in the acceptability rates [@WagersEtAl:2009]. 
When reading times are checked, they found that the participants read the ungrammatical sentences with a plural attractor easier then the ungrammatical sentences without a plural attractor. 
Similarly, participants found ungrammatical sentences with a plural attractor acceptable more often than that of with a singular attractor.

Some theories [@EberhardEtAl:2005; @HammerlyEtAl:2019] explain this phenomenon through a faulty representation of the number on the head of the agreement controller whereas other theories assume that agreement attraction is due to an error in the access to the number of the agreement controller [@WagersEtAl:2009; @NicenboimEtAl:2018].

Agreement attraction effects can equally well be explained by assuming shallow processing. Recent studies in psychology literature show that in certain environments people may not notice ungrammaticalities in a sentence or may not compute dependencies in a sentence completely [@OteroKintsch:1992; @FerreiraBaileyFerraro:2002].

To investigate this possibility, we report two speeded acceptability judgment experiments in this paper. The first one inspects the recent @LagoEtAl2018 findings. They found number attraction effects in genitive-possessive constructions such as `_[[the child's] toy]_' ([[_çocuğ-un\textsubscript{Gen}_] _oyuncağ-ı\textsubscript{Poss}_]). They justify the use of this construction with the findings of @NicolEtAl:2016. Since the genitive case is a possible subject marker, the participants can consider genitive marked attractors a candidate for the agreement controller. However, case marking on all subject heads is locally ambiguous the accusative and the possessive. In our experiment 1, we disambiguate the subject head and test whether or not the results were due to the shallow processing of the head subject and sticking with the accusative marked representation under some conditions.

@@

# Theories of agreement attraction

The difference between the theories that explain the agreement attraction phenomenon lies in the assumption regarding the origin of the illusion. Some theories [@BockEberhard:1993; @ViglioccoEtAl:1995; @Eberhard:1997; @FranckEtAl:2002; @BergenGibson:2012] explain this phenomenon by hypothesizing that the representations formed from the linguistic input is faulty. One theory in this group is Marking and Morphing Theory. MMT suggests that the subject phrase can hold a continuous number value from 1 (unambiguously plural) to -1 (unambiguously singular). This value is affected by the _percolation_ of the grammatical feature which can happen in any direction. The implementation of such models suggests that the final number of the phrase depends on the number morphology on the head and the elements within the phrase, and the syntactic proximity of these elements to the root node of the phrase.

Another type of model within this family is based on the idea that linguistic input is noisy, and it is accompanied by a priori linguistic knowledge [@BergenGibson:2012]. This account is an extension of the noisy channel model proposed by @Levy2008, which uses Bayesian inference to model the participants' inference given the word string. @PatsonHusband:2016 provide additional evidence for this type of analysis by showing that final interpretations of agreement attraction cases are not based on the actual string, meaning that the participants infer a plural-marked head noun whereas it is not marked by any plural inflection.

On the other side, some theories assume that agreement attraction is due to an error in the access to the number of the agreement controller [@WagersEtAl:2009; @NicenboimEtAl:2018]. Unlike MMT, these models do not assume an erroneous representation to form. Rather, the agreement attraction effects surface when there is a feature interference in the ungrammatical sentences like \@ref(item:wagers2009). This interference is triggered due to the partial match between the sought retrieval cues and the ones within the subject phrase. On some occasions, when there is no single item matching the necessary cues from the verb, number related matching cue from the attractor may satisfy the retrieval.

# Agreement Attraction in Turkish

Recently, @LagoEtAl2018 demonstrated comparable agreement attraction effects in comprehension with Turkish stimuli. 
They used the speeded acceptability judgment method with the following question that asks participants whether or not the sentence sounded acceptable. 
They used sentences like (\@ref(item:LagoExp))\footnote{Abbreviations used in the paper are as follows: \printglossaries} with genitive-marked attractors.
They manipulated the number morphology on the attractor (plural x singular) and the grammaticality of the verb (grammatical (SG) x ungrammatical (PL)). The example given here represents all four conditions.  
In an ungrammatical condition with a plural attractor, the word _ressamların_ (painters') is expected to create a grammaticality illusion and mistaken for the agreement controller instead of the head subject _rakibi_ (rival\textsubscript{poss}).

@@

Following previous work on agreement attraction, they
argue
that non-intervening genitive marked attractor may create grammaticality illusions in Turkish. They base their argument on the inhibition constraint proposed by @NicolEtAl:2016. @NicolEtAl:2016 show in a production experiment that genitive attractors do not trigger grammaticality illusions. According to their hypothesis, the reason behind their results is the fact that the participants may never consider genitive-marked NPs as a potential candidate for subjects.
However, this is not the case in Turkish. In embedded sentences, the definite subject is always marked with the genitive case, which makes the genitive-marked NPs a candidate for the subjecthood in Turkish.

However, a potential complication in interpreting these results is the fact that all head nouns in Lago et al.'s stimuli were morphologically ambiguous between possessive and accusative case. This ambiguity is the result of the consonant-ending subject head usage. In Turkish, both the accusative case and the possessive surface as -I. \footnote{Throughout the paper, we present the archiphonemes with the capitalized form of the symbol. The suffix -I can surface as \textit{-i, -ı, -u, -ü}. }
Because of this local ambiguity, all experimental sentences have two possible parses up until to the end of the sentence. Check the following examples in \@ref(parse:LagoExpAcc) and \@ref(parse:LagoExpPoss).

@@

The local ambiguity between the accusative and the possessive is not resolved until the verb _uzaklaştı_. Regardless of the assumption for parallel or serial processing, there may exist certain conditions where participants are stuck with a rival-\Acc{} representation. Since accusative number agreement controllers are extremely rare, while genitive agreement controllers are very frequent in Turkish, the ambiguity of the head noun may occasionally render the genitive modifier a stronger contender for being the agreement controller, resulting in agreement attraction. The problem with such cases for the Inhibition hypothesis is that it is not clear whether the possible non-subjecthood clue from the accusative case plays a role or the possible subjecthood clue from the genitive case plays a role.

# Experiment 1

```{r Exp1:DataWrangling, message=FALSE, warning=FALSE, include=FALSE, paged.print=FALSE}

fname_data <- "../workspace_exp1/exp_data.rds"
data_exp1 <- readRDS(fname_data)
fname_form <- "../workspace_exp1/exp_form.rds"
form_exp1 <- readRDS(file = fname_form)

### note: the low accuracy on fillers is not caused by a couple of items
###       -> see below
# filler_avgs <- 
# data_exp1 %>% subset(Type == "filler") %>% group_by(Item, condition) %>% 
#               summarize(M = mean(ResponseYes, na.rm = T)) %>%
#               arrange(condition, Item)
# fillers %>% subset(condition == "a") 
# fillers %>% subset(condition == "b")
###

# compute by-subject percentages of 'yes' responses, and average RTs 
avg_by_subj <- data_exp1 %>%
                group_by(subject, experiment, condition, 
                         grammatical, verb_num, attractor_num) %>%
                summarize(avRT = mean(RT), 
                          p_yes = mean(ResponseYes, na.rm = T), 
                          N = sum(!is.na(ResponseYes))  )

# reformat by-subject averages to a wide format
avg_by_subj_wide <- avg_by_subj %>% 
                      mutate(expcond = paste(experiment, condition, sep="_")) %>% 
                      ungroup() %>%
                      dplyr::select(-experiment, -condition, -avRT, -N,
                                    -grammatical, -verb_num, -attractor_num) %>%
                      tidyr::spread(expcond, p_yes) %>% 
                      mutate(delta_dc = AgrAttr_d - AgrAttr_c)

# Load Lago et al.'s monolingual data
fname_lagoetal <- "../Data/Lago_et_al/Lago_data.csv"
df_lagoetal <- read.csv(fname_lagoetal, encoding = "UTF-8", as.is = T)
df_lagoetal %<>% subset(Group == "monolingual")
df_lagoetal %<>% dplyr::select(-Accuracy, -L1:-Group, -List:-SelfRateGerman)

# Note: All rows with Experiment == "offline" also seem to be for 
#       the UNP task ('Grammatical' is NA). I wonder if these were
#       the same subjects, or if the subject labels were simply the same
#       for the two experiments.
with(df_lagoetal, stopifnot( is.na(Grammatical) == (Experiment == "offline") ))

df_lagoetal_unp <- df_lagoetal %>% 
                    subset(is.na(Grammatical)) %>%
                    dplyr::select(-Grammatical:-Label)
df_lagoetal_attr <- df_lagoetal %>% 
                  subset(!is.na(Grammatical)) %>%
                  dplyr::select(-Distance:-NewCond)

df_lagoetal_attr %<>% mutate(ResponseYes = (Response == "yes") ) %>% 
                      dplyr::select(-Response)
df_lagoetal_attr %<>% ungroup() %>%
                      dplyr::select(grammatical=Grammatical,
                                    attractor_num=Attractor,
                                    experiment=Experiment,
                                    lagoetal_condition=Condition, 
                                    subject=Participant, 
                                    item=Item,
                                    ResponseYes,
                                    RT)

# map to our condition labels
lagoetal_condition_mapping <- data.frame( condition = c("a", "b", "c", "d"),
                                          lagoetal_condition = c("d", "b", "c", "a"), 
                                          stringsAsFactors = F)
df_lagoetal_attr %<>% left_join( lagoetal_condition_mapping, by = "lagoetal_condition" )

# compute by-subject percentages of 'yes' responses, and average RTs 
avg_by_subj_lagoetal <- df_lagoetal_attr %>%
                            group_by(subject, experiment, condition, grammatical, attractor_num) %>%
                            summarize(avRT = mean(RT), 
                                      p_yes = mean(ResponseYes, na.rm = T), 
                                      N = sum(!is.na(ResponseYes))  )

# reformat by-subject averages to a wide format
avg_by_subj_lagoetal_wide <- avg_by_subj_lagoetal %>% 
                                    mutate(expcond = paste(experiment, condition, sep="_")) %>% 
                                    ungroup() %>%
                                    dplyr::select(-experiment, -condition, -avRT, -N,
                                                  -grammatical, -attractor_num) %>%
                                    tidyr::spread(expcond, p_yes) %>% 
                                    mutate(delta_dc = online_d - online_c)

# identify bad participants 
bad_subjects <- subset(avg_by_subj_wide, delta_dc <= 0.25 ) %>% .$subject
data_exp1_clean <- data_exp1 %>% subset(!subject %in% bad_subjects)

# identify bad participants 
bad_subjects_lagoetal <- subset(avg_by_subj_lagoetal_wide, delta_dc <= 0.25 ) %>% .$subject
df_lagoetal_attr_clean <- df_lagoetal_attr %>% subset(!subject %in% bad_subjects_lagoetal)

# merge both datasets
df_merge_exp1 <- data_exp1_clean %>% ungroup() %>% 
                      dplyr::select(source=experiment, 
                                    grammatical, 
                                    attractor_num,
                                    # condition,
                                    subject, 
                                    item=Item,
                                    ResponseYes, 
                                    RT)
df_merge_exp1$experiment <- "Experiment 1"
df_merge_exp1$grammatical %<>% dplyr::recode(gram="grammatical", ungram="ungrammatical")
df_merge_exp1$attractor_num %<>% dplyr::recode(pl="plural", sg="singular")

df_merge_exp1$item %<>% as.integer()
df_merge_exp1$subject %<>% as.character()

df_merge_lago <- df_lagoetal_attr_clean %>%
                      ungroup() %>% 
                      dplyr::select(grammatical, attractor_num,
                                    subject, item, ResponseYes, RT)
df_merge_lago$experiment <- "Lago et al. (2018)" 
df_merge_lago$source <- NA

```

```{r Exp1:wordFreq, message=FALSE, warning=FALSE, include=FALSE, paged.print=FALSE}
# word frequencies: 
# Utku: are these still relevant for us?

word_freq <- readxl::read_excel("../Data/frequencies_exp1_and_lago.xlsx", sheet = 1)
word_freq$freq_percentage %<>% as.numeric()
word_freq %<>% dplyr::select(-freq_standardized, -freq_percentage, -word)
word_freq %<>% tidyr::spread(place, freq_count)

word_freq %<>% dplyr::group_by(exp) %>% 
               dplyr::mutate( freq_cat_n1 = ifelse(n1 > median(n1), "high", "low"),
                              freq_cat_n2 = ifelse(n2 > median(n2), "high", "low"),
                              freqlog_n1 = scale(log(n1)),
                              freqlog_n2 = scale(log(n2)),
                              freqlog_n1n2 = log(n1) - log(n2)
                            ) %>% 
                ungroup()
# word_freq$freq_cat_n1 %<>% as.factor()
# word_freq$freq_cat_n2 %<>% as.factor()

word_freq_exp1 <- word_freq %>% subset(exp == "exp1") %>% dplyr::select(-exp)
df_merge_exp1 %<>% left_join(word_freq_exp1, by = "item")

word_freq_lago <- word_freq %>% subset(exp == "lagoetal") %>% dplyr::select(-exp)
df_merge_lago %<>% left_join(word_freq_lago, by = "item")
df_merge_lago$item %<>% add(1000)

df_merged <- dplyr::bind_rows(df_merge_exp1, df_merge_lago)
df_merged$subject %<>% as.factor()
df_merged$item %<>% as.factor()
```

Previous work in agreement attraction found significant effects in comprehension using forced-choice acceptability judgment tasks [@HammerlyEtAl:2019; @LagoEtAl2018; @WagersEtAl:2009]. The goal of experiment 1 is to control for the effect of head noun ambiguity on the magnitude of agreement attraction effect. To this end, we replicated @LagoEtAl2018 experiment with unambiguous head nouns. We disambiguated the cases by using vowel-ending head nouns. When attached to a vowel-ending noun, the accusative case and the possessive surface with a different floating consonant. For example, we changed sentence (\@ref(item:LagoExpContrast)) to sentence (\@ref(item:OurExpContrast)).

@@

## Participants

We recruited 118 undergraduate students to participate in Experiment 1 in exchange for course credit. All participants were native speakers of Turkish, with an average age of `r printnum(mean(form_exp1$Age))` (range: `r min(form_exp1$Age)` -- `r max(form_exp1$Age)`).  
Experiment were carried out following the Declaration of Helsinki and the regulations concerning ethics at research in Bo\u{g}azi\c{c}i University. All participants provided informed consent prior to their participation.

## Materials

>>>>

% Options for packages loaded elsewhere
\PassOptionsToPackage{unicode}{hyperref}
\PassOptionsToPackage{hyphens}{url}
%
\documentclass[
  english,
  doc,floatsintext]{apa6}

\usepackage{graphicx}
\makeatletter
\def\maxwidth{\ifdim\Gin@nat@width>\linewidth\linewidth\else\Gin@nat@width\fi}
\def\maxheight{\ifdim\Gin@nat@height>\textheight\textheight\else\Gin@nat@height\fi}
\makeatother
% Scale images if necessary, so that they will not overflow the page
% margins by default, and it is still possible to overwrite the defaults
% using explicit options in \includegraphics[width, height, ...]{}
\setkeys{Gin}{width=\maxwidth,height=\maxheight,keepaspectratio}
% Set default figure placement to htbp
\makeatletter
\def\fps@figure{htbp}
\makeatother
\setlength{\emergencystretch}{3em} % prevent overfull lines
\providecommand{\tightlist}{%
  \setlength{\itemsep}{0pt}\setlength{\parskip}{0pt}}
\setcounter{secnumdepth}{5}
% Make \paragraph and \subparagraph free-standing
\ifx\paragraph\undefined\else
  \let\oldparagraph\paragraph
  \renewcommand{\paragraph}[1]{\oldparagraph{#1}\mbox{}}
\fi
\ifx\subparagraph\undefined\else
  \let\oldsubparagraph\subparagraph
  \renewcommand{\subparagraph}[1]{\oldsubparagraph{#1}\mbox{}}
\fi

%\usepackage[utf8x]{inputenc}
\usepackage{amsmath}
\usepackage{amsthm}
\usepackage{graphicx}
\usepackage[colorinlistoftodos]{todonotes}
\usepackage[singlespacing]{setspace}

\let\B\relax

\let\T\relax

\usepackage{gb4e}
\exewidth{(\thexnumi)} %ex number no indent setting
\let\eachwordone=\textit %first line italic
\noautomath

\usepackage[linguistics,edges]{forest} 
\forestset{
ned/.style={%
    for tree={calign=fixed edge angles},
    delay={%
      where content={}{%
        shape=coordinate,
        for nodewalk={%
          Nodewalk={%
            on invalid=fake,
          }{%
            parent,
          }{%
            for children={anchor=north},
          }
        }{},
      }{},
    },
},
fned/.style={
        delay={where content={}{shape=coordinate,
              for siblings={anchor=north}}{}},
		for tree={s sep=4mm}},
}

\usepackage{leipzig}
\makeglossaries

\usepackage{hyperref}
\hypersetup{colorlinks = true, linkcolor = blue, citecolor = blue, urlcolor = blue}
\usepackage{multicol}
% Manuscript styling
\usepackage{upgreek}
\captionsetup{font=singlespacing,justification=justified}

% Table formatting
\usepackage{longtable}
\usepackage{lscape}
% \usepackage[counterclockwise]{rotating}   % Landscape page setup for large tables
\usepackage{multirow}		% Table styling
\usepackage{tabularx}		% Control Column width
\usepackage[flushleft]{threeparttable}	% Allows for three part tables with a specified notes section
\usepackage{threeparttablex}            % Lets threeparttable work with longtable

% Create new environments so endfloat can handle them
% \newenvironment{ltable}
%   {\begin{landscape}\begin{center}\begin{threeparttable}}
%   {\end{threeparttable}\end{center}\end{landscape}}
\newenvironment{lltable}{\begin{landscape}\begin{center}\begin{ThreePartTable}}{\end{ThreePartTable}\end{center}\end{landscape}}

% Enables adjusting longtable caption width to table width
% Solution found at http://golatex.de/longtable-mit-caption-so-breit-wie-die-tabelle-t15767.html
\makeatletter
\newcommand\LastLTentrywidth{1em}
\newlength\longtablewidth
\setlength{\longtablewidth}{1in}
\newcommand{\getlongtablewidth}{\begingroup \ifcsname LT@\roman{LT@tables}\endcsname \global\longtablewidth=0pt \renewcommand{\LT@entry}[2]{\global\advance\longtablewidth by ##2\relax\gdef\LastLTentrywidth{##2}}\@nameuse{LT@\roman{LT@tables}} \fi \endgroup}

% \setlength{\parindent}{0.5in}
% \setlength{\parskip}{0pt plus 0pt minus 0pt}

% \usepackage{etoolbox}
\makeatletter
\patchcmd{\HyOrg@maketitle}
  {\section{\normalfont\normalsize\abstractname}}
  {\section*{\normalfont\normalsize\abstractname}}
  {}{\typeout{Failed to patch abstract.}}
\patchcmd{\HyOrg@maketitle}
  {\section{\protect\normalfont{\@title}}}
  {\section*{\protect\normalfont{\@title}}}
  {}{\typeout{Failed to patch title.}}
\makeatother
\shorttitle{Agreement Attraction in Turkish}
\keywords{keywords\newline\indent Word count: X}
\usepackage{csquotes}
\geometry{left=25mm, right=25mm, top=25mm, bottom=25mm}

\ifxetex
  % Load polyglossia as late as possible: uses bidi with RTL langages (e.g. Hebrew, Arabic)
  \usepackage{polyglossia}
  \setmainlanguage[]{english}
\else
  \usepackage[shorthands=off,main=english]{babel}
\fi
\ifluatex
  \usepackage{selnolig}  % disable illegal ligatures
\fi
\newlength{\cslhangindent}
\setlength{\cslhangindent}{1.5em}
\newenvironment{cslreferences}%
  {\setlength{\parindent}{0pt}%
  \everypar{\setlength{\hangindent}{\cslhangindent}}\ignorespaces}%
  {\par}

\title{Agreement Attraction in Turkish}
\author{Utku Türk\textsuperscript{1}}
\date{}

\authornote{

authornote

Correspondence concerning this article should be addressed to Utku Türk, Postal address. E-mail: \href{mailto:utku.turk@boun.edu.tr}{\nolinkurl{utku.turk@boun.edu.tr}}

}

\affiliation{\vspace{0.5cm}\textsuperscript{1} Boğaziçi University University, Istanbul, Turkey}

\abstract{
We report the results of two speeded acceptability judgment experiments in Turkish. We hypothesized an alternative explanation for agreement attraction effects in Turkish that is based on shallow processing. Our findings contradict our hypothesized form-driven processing strategy and support an account of agreement attraction based on the use of abstract linguistic features, rather than mere form.
}

\begin{document}
\maketitle

\hypertarget{introduction}{%
\section{Introduction}\label{introduction}}

It has been observed that speakers often fail to accurately process grammatical dependencies between the parts of a sentence (Gibson \& Thomas, 1999; Phillips, Wagers, \& Lau, 2011).
For example, ungrammatical center-embedded structures like \emph{The patient who the nurse who the clinic had hired met Jack} are found grammatical by native English speakers.
One such comprehension error is agreement attraction (Nicol, Forster, \& Veres, 1997; Pearlmutter, Garnsey, \& Bock, 1999; Wagers, Lau, \& Phillips, 2009).
Speakers may erroneously find sentences acceptable in which the verb erroneously agrees with a syntactically unrelated noun phrase (\emph{the attractor}) instead of the subject (\emph{the agreement controller}) as in \ref{item:wagers2009}.

@@

These effects were found to be robust in a variety of constructions in different languages, e.g.~English , German , Russian, Arabic, Spanish, Armenian, Slovak, Greek, and Turkish (Avetisyan, Lago, \& Vasishth, 2020; Badecker \& Kuminiak, 2007; Lago et al., 2018; Lago, Shalom, Sigman, Lau, \& Phillips, 2015; Paspali \& Marinis, 2020; Reifegerste, Hauer, \& Felser, 2016; Tucker, Idrissi, \& Almeida, 2015) using structures like object relative clauses, subject relative clauses, possessive relative clauses, and genitive-possessive noun phrases (Dillon, Clifton, \& Frazier, 2013; Franck, Colonna, \& Rizzi, 2015; Häussler \& Bader, 2009; Lago et al., 2018, 2015; Parker, Lago, \& Phillips, 2015; Tucker et al., 2015; Wagers et al., 2009). These effects are found both in grammatical marking of the gender and the number (Badecker \& Kuminiak, 2007; Paspali \& Marinis, 2020; Slioussar \& Malko, 2016), and attested with various methods including maze tasks, ERP, eye tracking to name a few (Nicol et al., 1997; Tanner, Nicol, \& Brehm, 2014; Wagers et al., 2009).

One instance of agreement attraction is found in the production processes in which participants were asked to complete sentence fragments (Bock \& Miller, 1991). Participants used the wrong verb more often when the attractor (\emph{cabinets}) were plural, and the agreement controller (\emph{key}) were singular as in \ref{item:bock1991}.

@@

Agreement attraction is also found in the comprehension studies. In these studies, agreement attraction is found either as a facilitation in reading times or as a effect in the acceptability rates (Wagers et al., 2009).
When reading times are checked, they found that the participants read the ungrammatical sentences with a plural attractor easier then the ungrammatical sentences without a plural attractor.
Similarly, participants found ungrammatical sentences with a plural attractor acceptable more often than that of with a singular attractor.

Some theories (Eberhard, Cutting, \& Bock, 2005; Hammerly, Staub, \& Dillon, 2019) explain this phenomenon through a faulty representation of the number on the head of the agreement controller whereas other theories assume that agreement attraction is due to an error in the access to the number of the agreement controller (Nicenboim, Vasishth, Engelmann, \& Suckow, 2018; Wagers et al., 2009).

Agreement attraction effects can equally well be explained by assuming shallow processing. Recent studies in psychology literature show that in certain environments people may not notice ungrammaticalities in a sentence or may not compute dependencies in a sentence completely (Ferreira, Bailey, \& Ferraro, 2002; Otero \& Kintsch, 1992).

To investigate this possibility, we report two speeded acceptability judgment experiments in this paper. The first one inspects the recent Lago et al. (2018) findings. They found number attraction effects in genitive-possessive constructions such as `\emph{{[}{[}the child's{]} toy{]}}' ({[}{[}\emph{çocuğ-un\textsubscript{Gen}}{]} \emph{oyuncağ-ı\textsubscript{Poss}}{]}). They justify the use of this construction with the findings of Nicol, Barss, and Barker (2016). Since the genitive case is a possible subject marker, the participants can consider genitive marked attractors a candidate for the agreement controller. However, case marking on all subject heads is locally ambiguous the accusative and the possessive. In our experiment 1, we disambiguate the subject head and test whether or not the results were due to the shallow processing of the head subject and sticking with the accusative marked representation under some conditions.

@@

\hypertarget{theories-of-agreement-attraction}{%
\section{Theories of agreement attraction}\label{theories-of-agreement-attraction}}

The difference between the theories that explain the agreement attraction phenomenon lies in the assumption regarding the origin of the illusion. Some theories (Bergen \& Gibson, 2012; Bock \& Eberhard, 1993; Eberhard, 1997; Franck, Vigliocco, \& Nicol, 2002; Vigliocco, Butterworth, \& Semenza, 1995) explain this phenomenon by hypothesizing that the representations formed from the linguistic input is faulty. One theory in this group is Marking and Morphing Theory. MMT suggests that the subject phrase can hold a continuous number value from 1 (unambiguously plural) to -1 (unambiguously singular). This value is affected by the \emph{percolation} of the grammatical feature which can happen in any direction. The implementation of such models suggests that the final number of the phrase depends on the number morphology on the head and the elements within the phrase, and the syntactic proximity of these elements to the root node of the phrase.

Another type of model within this family is based on the idea that linguistic input is noisy, and it is accompanied by a priori linguistic knowledge (Bergen \& Gibson, 2012). This account is an extension of the noisy channel model proposed by Levy (2008), which uses Bayesian inference to model the participants' inference given the word string. Patson and Husband (2016) provide additional evidence for this type of analysis by showing that final interpretations of agreement attraction cases are not based on the actual string, meaning that the participants infer a plural-marked head noun whereas it is not marked by any plural inflection.

On the other side, some theories assume that agreement attraction is due to an error in the access to the number of the agreement controller (Nicenboim et al., 2018; Wagers et al., 2009). Unlike MMT, these models do not assume an erroneous representation to form. Rather, the agreement attraction effects surface when there is a feature interference in the ungrammatical sentences like \ref{item:wagers2009}. This interference is triggered due to the partial match between the sought retrieval cues and the ones within the subject phrase. On some occasions, when there is no single item matching the necessary cues from the verb, number related matching cue from the attractor may satisfy the retrieval.

\hypertarget{agreement-attraction-in-turkish}{%
\section{Agreement Attraction in Turkish}\label{agreement-attraction-in-turkish}}

Recently, Lago et al. (2018) demonstrated comparable agreement attraction effects in comprehension with Turkish stimuli.
They used the speeded acceptability judgment method with the following question that asks participants whether or not the sentence sounded acceptable.
They used sentences like (\ref{item:LagoExp})\footnote{Abbreviations used in the paper are as follows: \printglossaries} with genitive-marked attractors.
They manipulated the number morphology on the attractor (plural x singular) and the grammaticality of the verb (grammatical (SG) x ungrammatical (PL)). The example given here represents all four conditions.\\
In an ungrammatical condition with a plural attractor, the word \emph{ressamların} (painters') is expected to create a grammaticality illusion and mistaken for the agreement controller instead of the head subject \emph{rakibi} (rival\textsubscript{poss}).

@@

Following previous work on agreement attraction, they
argue
that non-intervening genitive marked attractor may create grammaticality illusions in Turkish. They base their argument on the inhibition constraint proposed by Nicol et al. (2016). Nicol et al. (2016) show in a production experiment that genitive attractors do not trigger grammaticality illusions. According to their hypothesis, the reason behind their results is the fact that the participants may never consider genitive-marked NPs as a potential candidate for subjects.
However, this is not the case in Turkish. In embedded sentences, the definite subject is always marked with the genitive case, which makes the genitive-marked NPs a candidate for the subjecthood in Turkish.

However, a potential complication in interpreting these results is the fact that all head nouns in Lago et al.'s stimuli were morphologically ambiguous between possessive and accusative case. This ambiguity is the result of the consonant-ending subject head usage. In Turkish, both the accusative case and the possessive surface as -I. \footnote{Throughout the paper, we present the archiphonemes with the capitalized form of the symbol. The suffix -I can surface as \textit{-i, -ı, -u, -ü}. }
Because of this local ambiguity, all experimental sentences have two possible parses up until to the end of the sentence. Check the following examples in \ref{parse:LagoExpAcc} and \ref{parse:LagoExpPoss}.

@@

The local ambiguity between the accusative and the possessive is not resolved until the verb \emph{uzaklaştı}. Regardless of the assumption for parallel or serial processing, there may exist certain conditions where participants are stuck with a rival-\Acc{} representation. Since accusative number agreement controllers are extremely rare, while genitive agreement controllers are very frequent in Turkish, the ambiguity of the head noun may occasionally render the genitive modifier a stronger contender for being the agreement controller, resulting in agreement attraction. The problem with such cases for the Inhibition hypothesis is that it is not clear whether the possible non-subjecthood clue from the accusative case plays a role or the possible subjecthood clue from the genitive case plays a role.

\hypertarget{experiment-1}{%
\section{Experiment 1}\label{experiment-1}}

Previous work in agreement attraction found significant effects in comprehension using forced-choice acceptability judgment tasks (Hammerly et al., 2019; Lago et al., 2018; Wagers et al., 2009). The goal of experiment 1 is to control for the effect of head noun ambiguity on the magnitude of agreement attraction effect. To this end, we replicated Lago et al. (2018) experiment with unambiguous head nouns. We disambiguated the cases by using vowel-ending head nouns. When attached to a vowel-ending noun, the accusative case and the possessive surface with a different floating consonant. For example, we changed sentence (\ref{item:LagoExpContrast}) to sentence (\ref{item:OurExpContrast}).

@@

\hypertarget{participants}{%
\subsection{Participants}\label{participants}}

We recruited 118 undergraduate students to participate in Experiment 1 in exchange for course credit. All participants were native speakers of Turkish, with an average age of 20.14 (range: 18 -- 32).\\
Experiment were carried out following the Declaration of Helsinki and the regulations concerning ethics at research in Bo\u{g}azi\c{c}i University. All participants provided informed consent prior to their participation.

\hypertarget{materials}{%
\subsection{Materials}\label{materials}}

We used 40 sets of sentences like \ref{item:exp1ExperimentalItems}, in which we manipulated
(i) the number of the attractor noun, and
(ii) the number agreement on the verb.
Plural number and plural agreement were both marked with the affix \textit{-ler/-lar}, while singular number and singular agreement were marked with its absence.
We have used the experimental items in Lago et al. (2018) as a base.
All sentences started with a complex subject NP like \textit{`the manager's cook'} (\textit{yöneticinin aşcısı}), in which the possessor, which functioned as the attractor, carried genitive case, and the head noun carried unambiguous possessive case.
Because the head noun was singular in all conditions, sentences with plural verb agreement were ungrammatical.
Moreover, as in the original study, the properties of the nouns in the complex subject NP and the relation between these nouns held constant in our replication.
The distribution of the verb types followed the same pattern as in Lago et al. (2018), namely twenty unergatives, eighteen unaccusatives, and two optionally transitive verbs.
Both in Lago et al. (2018) and in our replication, pre-verbal adverbials consisting of 2-3 words (15 characters on average) are used. We changed this adverbials for plausability reasons, such that with head nouns like cooks we used locational adverbs like kitchen, instead of the stage.
Unlike the Lago et al. (2018) sentences, our head nouns were vowel-ending, and therefore not ambiguous between accusative case, for which the case suffix is \textit{-yI}, and possessive case, for which the suffix is \textit{-sI}.
